---
title: "esm206-lab5-cje"
author: "Chance English"
date: "10/25/2021"
output: 
  html_document: 
    toc: yes
    theme: united
    code_folding: hide
---

```{r setup, include = T, message = F, warning = F}
knitr::opts_chunk$set(echo = TRUE, include = T, message = F, warning = F)

library(tidyverse)
library(effsize)
```

## Setup & change RMarkdown settings

Here, we'll use the "repo first" strategy to create a version-controlled R project. 

- Make a new repo on GitHub (called esm206-lab5 or similar), with a ReadME
- Clone to create a version-controlled R project with the same name

We'll use the 'Settings > Output Settings' tool to update the YAML, changing the theme, highlighting, and adding a table of contents. 

- Create a new R Markdown document, don't delete anything
- Go to the gear icon "Settings" dropdown at the top of your .Rmd, and choose Output Options, where you can change the theme, add a table of contents, and more

### Confidence intervals 

To find a confidence interval with a single sample, use `t.test()` on the vector of sample values:
 
 
```{r}
# Create my sample (vector):
otter_length <- c(38, 41, 50, 27, 32, 41, 48, 60, 43)

# Use t.test() to return a CI (95% is default)
otter_ci <- t.test(otter_length)

### default tests that the mean otter length is 0 which is what the t.test will use to test (this is a one sample t-test)

# CI: 34.69 - 49.75. If we took multiple samples from this population, and calculated a CI for each of them then we expect 95% of those samples to contain the true mean. This is a range which probably contains the true population mean

# Find the sd
otter_sd <- sd(otter_length)
```

Mean otter length is 42.2 $\pm$ 9.3 inches (mean $\pm$ sd,n = 8) with a 95% CI of otter_ci$conf.int inches. 

#### Now, using inline referencing 

Mean otter length `r round(otter_ci$estimate, 1)` inches 

