---
title: "esm206-lab5-cje"
author: "Chance English"
date: "10/25/2021"
output: 
  html_document: 
    toc: yes
    theme: united
    code_folding: hide
---

```{r setup, include = T, message = F, warning = F}
knitr::opts_chunk$set(echo = TRUE, include = T, message = F, warning = F)

library(tidyverse)
library(effsize)
```

## Setup & change RMarkdown settings

Here, we'll use the "repo first" strategy to create a version-controlled R project. 

- Make a new repo on GitHub (called esm206-lab5 or similar), with a ReadME
- Clone to create a version-controlled R project with the same name

We'll use the 'Settings > Output Settings' tool to update the YAML, changing the theme, highlighting, and adding a table of contents. 

- Create a new R Markdown document, don't delete anything
- Go to the gear icon "Settings" dropdown at the top of your .Rmd, and choose Output Options, where you can change the theme, add a table of contents, and more

## Part 1: Confidence intervals

### Example 1

To find a confidence interval with a single sample, use `t.test()` on the vector of sample values:
 
 
```{r}
# Create my sample (vector):
otter_length <- c(38, 41, 50, 27, 32, 41, 48, 60, 43, 90, 72, 112)

# Use t.test() to return a CI (95% is default)
otter_ci <- t.test(otter_length)

### default tests that the mean otter length is 0 which is what the t.test will use to test (this is a one sample t-test)

# CI: 34.69 - 49.75. If we took multiple samples from this population, and calculated a CI for each of them then we expect 95% of those samples to contain the true mean. This is a range which probably contains the true population mean

# Find the sd
otter_sd <- sd(otter_length)
```


### Example 2: Inline referencing 

Mean otter length is `r round(otter_ci$estimate, 1)` inches with a 95% confidence interval of [`r round(otter_ci$conf.int[1], 1)`, `r round(otter_ci$conf.int[2],1)`] inches.


## Part 2: Two Sample t-test

### Example 1

Make two mock samples, and ask the question: do we have sufficient evidence to reject the null hypothesis that the samples were drawn from populations with the same mean



```{r}

oak_trees <- c(20,19,24,28,43,12,15,20, 200, 180, 300)


pine_trees <- c(39,26,20,30,37,22,19,18)

# Two Sample t-test

trees_t <- t.test(oak_trees,pine_trees)



```

P = 0.4149. If our samples were drawn from the same population, there is a 42% chance of taking a random samples that are at least as different as the random samples we've observed here. 

The mean heights of oak and pine were (`r round(mean(oak_trees),1)` $\pm$ `r round(sd(oak_trees),1)`) and (`r round(mean(pine_trees),1)` $\pm$ `r round(sd(pine_trees),1)`), respectively and the difference was not significant (t(`r round(trees_t$parameter, 1)`) = `r round(trees_t$statistic, 1)`, p = `r round(trees_t$p.value,1)`).


## Part 3: Another t-test using mpg data

### Example 1: Mpg subset for suv 

```{r}

compact_suv <- mpg %>% 
  filter(class %in% c("compact", "suv"))


# Confirm Subset 

unique(compact_suv$class)

```

Create a histogram of city gas mileage faceted by class starting from compact_suv 

```{r}

compact_suv %>% 
  ggplot(aes(x = cty)) +
  geom_histogram(bins = 12) +
  facet_wrap(~class)

ggplot(data = compact_suv, aes(sample = cty)) +
  geom_qq() +
  facet_wrap(~class)



```


Find some summary statistics for each group 

```{r}

car_stats <- compact_suv %>% 
  group_by(class) %>% 
  summarize(
    mean_city = round(mean(cty), 2),
    sd_city = round(sd(cty),2),
    sample_size = n()
  )

car_stats

```

The t-test function really wants vectors. Use the pull() to pull out vectros from dataframe 

```{r}

compact_sample <- compact_suv %>% 
  filter(class == "compact") %>% 
  pull(cty)


suv_sample <- compact_suv %>% 
  filter(class == "suv") %>% 
  pull(cty)


```


```{r}

compact_suv_t <- t.test(compact_sample, suv_sample)
compact_suv_t


```


The mean mpg of compact and suv cars were (`r round(mean(compact_sample),1)` $\pm$ `r round(sd(compact_sample),1)`) and (`r round(mean(suv_sample),1)` $\pm$ `r round(sd(suv_sample),1)`) mpg, respectively and the difference was significant (t(`r round(compact_suv_t$parameter)`) = `r round(compact_suv_t$statistic, 1)`, p = `r round(compact_suv_t$p.value,1)`).
